---
title: "Scraping"
author: "Aubrey DuBois"
date: '2022-07-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#load packages
```{r load packages}

library(pdftools)
library(tidyverse)
library(reticulate)
library(rscala)
#library(tm)
#library(lubridate)
#library(kableExtra)
#library(tidytext)
#library(quanteda)
#library(textreadr)


```

#read in pdfs and get basic paper info

```{r read pdfs and get basic paper info}

#creates a list of the pdf files

### FOR MAC/LINUX ###
#system function allows you to run command line functions
pdfs <- system("ls -1 PDFs/*.pdf", intern = T)

### FOR WINDOWS ###
#create a list of all the PDF documents
#pdfs <- dir("C:/Users/duboisa7/Documents/GitKraken/BB/Blueberry/PDFs",
#             full.names = T,
#             recursive = T,
#             pattern = "*.pdf") %>%
#  as.list()



#create a function
#create a df called pdf.info to store results of pdf_info function
#if statements ensure that if file is missing data, it will list NA in that column
#return keeps only the specified columns

fun <- function(file){
  
  result <- pdf_info(file) %>% 
    data.frame()
  
  if(!("keys.doi" %in% colnames(result))){
    result$keys.doi <- "NA"
  }
  if(!("keys.Author" %in% colnames(result))){
    result$keys.Author <- "NA"
  }
  if(!("keys.Title" %in% colnames(result))){
    result$keys.Title <- "NA"
  }
  if(!("created" %in% colnames(result))){
    result$created <- "NA"
  }
  
  return(result[,c("keys.Title","keys.Author", "created", "keys.doi")]) 
}



## USING LAPPLY ##
#create a temporary object that uses lapply to apply the created 'fun' function to all pdfs in the list 'pdfs'
temp <- lapply(pdfs, FUN=fun)

#create a df that binds each temp object to itself
pdf.info <- do.call(rbind, temp) %>%
  rename(DOI = keys.doi,
         Author = keys.Author,
         Title = keys.Title,
         Published = created)



## USING FOR LOOPS ##
#for each file in the pdfs list, apply the 'fun' function to the file and bind the results together
#for(file in pdfs){
#  loopresult <- rbind(loopresult, fun(file))
#}

```

```{r pull tables}




```


```{r pull materials and methods section}



```{r testing random methods}



#############################################################
#############################################################
#############################################################

r <- raw2 %>%
  as_tibble(text = text)


r %>%
  unnest_tokens(word, text)

key <-  "d35db8b36f9daa0bf919e69e06177c7d"


#create function 
get_data <- function(dat) {
    list(Full_Text = str_trim(str_extract(dat, "(?<=FULL TEXT)(?s)(.*?)(?=CREDIT)")),
       Pub_Title = str_trim(str_extract(dat, "(?<=Publication title:)(.*?)(?=;)")),
       Pub_Date = str_trim(str_extract(dat, "(?<=Publication date:)(.*?)(?=\\\\n)")),
       Pub_Language = str_trim(str_extract(dat, "(?<=Language of publication:)(.*?)(?=\\\\n)"))
  )
}



####################################################################################
####################################################################################
####################################################################################

t1 <- ft_get('10.1371/journal.pone.0138494', type = "xml")

t1 %>%
  pub_chunks(sections = "authors")

Sys.setenv(ELSEVIER_TDM_KEY = "d35db8b36f9daa0bf919e69e06177c7d")

t2 <- ft_get(x = "10.1016/j.jfca.2022.104412", from = "elsevier", type = "xml", elsevieropts = list(retain_non_ft = TRUE))




```
  